---
title: 'Assignment 3: Relationships'
author: "Julia Meinhardt"
date: "9/20/2020"
output: 
    html_document:
      toc: true
      toc_float: true
      highlight:  tango
      code_folding: hide
---


```{r setup, include=FALSE}
library(ggplot2)
library(tidyverse)
```


```{r load data, message=FALSE}
bay_data <- read_csv("baytractdata2018.csv")
jefferson_data <- read_csv("jeffersontractdata2018.csv")
combined_data <- read_csv("jeffersonbay2018.csv")
```

## Topic and Variables

My data analysis uses tract-level data to explore poverty, public assistance, and high-school graduation rates in two regions of California: the Bay Area and an informal region called "Jefferson" comprised of rural counties in Northeast California. Jefferson is a primarily rural region, while the Bay Area is primarily urban.

I have quite a large set of variables, most of which are continuous. Part of this is because I've chosen to look at all three poverty variables as defined by the 5-Year ACS as well as differences in public assistance including and excluding Social Security Income. 

**List of Variables**


1. Total Population: total_popE =  (continuous)
2. Total below 150 % poverty level: low_income (continuous)
3. Percent of population receiving public assistance (including SSI):  (continuous)
4. Percent of population receiving cash public assistance (including food stamps/SNAP):  (continuous)
5. High School Graduates:  (categorical, >83% meets statewide average)
6. Some college or Associates: (categorical, >29% meets statewide average)
7. Bachelors or higher: (categorical, >33% meets statewide average)
8. County (categorical)
9. Region (categorical)

 
#### Organization and Credit

Because I've chosen a large number of continuous variables and per the assignment instructions to test every possible relationship, there are quite a few calculations in this analysis. To make the document readable and emphasize salient information, my code chunks are all collapsible. If the reader would like to view all the code, you can use the code button at the top right corner of the screen to turn all code on or off. Additionally, you can view individual code chunks as desired by clicking the code button above the output for which you'd like to see the code. 

Tabs are organized by the type of test being performed. Within each tab, calculations are organized by whether or not they are being performed on the <span style="color: blue;">Bay Area Region</span> data, <span style="color: green;">Jefferson Region</span> data, or <span style="color: orange;">both regions combined</span>. 

The relationships between all variables as shown by the tests are all described in the **Relationship Strengths and Statistical Significance** tab. The calculations I've chosen to visualize and analyze are in the **Discussion, Visualizations, and Comparisons** tab. 

I used **Cat's** example and **Carole's** tutorial as the basis for organization and exploration. **Dylan**  shared a code function with me that allowed me to visualize my correlations in one table, which was extremely useful for visualizing correlations in one space. **Mel** shared the code for creating a floating table of contents in our Tuesday class meeting, which ended up being so essential for organizing this analysis in a way that was manageable. And **Ryan** reminded me that this was assignment 3 and not 2... Thanks all! 

## Tests, Analysis, and Visualizations{.tabset .tabset-fade .tabset-pills}

### Pearson's Correlation Calculations 


#### <span style="color: blue;">Bay Area Correlations Summary Table:</span>  
```{r bay correlation table}
cor(bay_data[,unlist(lapply(bay_data, is.numeric))])
```


#### <span style="color: green;">Jefferson Correlations Summary Table:</span>  
```{r jefferson correlation table}
cor(jefferson_data[,unlist(lapply(jefferson_data, is.numeric))])
```



The remainder of this tab shows individual correlations in more detail. These will be explored further in the **Visualizations, Comparisons, and Analysis** tab. 

#### <span style="color: blue;">Individual Bay Correlations</span>  
```{r individual correlations for bay}

correlation1_bay <- cor.test(bay_data$total_popE, bay_data$low_income)
correlation1_bay
correlation2_bay <- cor.test(bay_data$total_popE, bay_data$pub_assist)
correlation2_bay
correlation3_bay <- cor.test(bay_data$total_popE, bay_data$cash_assist)
correlation3_bay
correlation4_bay <- cor.test(bay_data$low_income, bay_data$pub_assist)
correlation4_bay
correlation5_bay <- cor.test(bay_data$low_income, bay_data$cash_assist)
correlation5_bay
correlation6_bay <- cor.test(bay_data$cash_assist, bay_data$pub_assist)
correlation6_bay
```

#### <span style="color: green;">Individual Jefferson Correlations</span>  

```{r individual correlations for Jefferson}
correlation1_jefferson <- cor.test(jefferson_data$total_popE, jefferson_data$low_income)
correlation1_jefferson
correlation2_jefferson <- cor.test(jefferson_data$total_popE, jefferson_data$pub_assist)
correlation2_jefferson
correlation3_jefferson <- cor.test(jefferson_data$total_popE, jefferson_data$cash_assist)
correlation3_jefferson
correlation4_jefferson <- cor.test(jefferson_data$low_income, jefferson_data$pub_assist)
correlation4_jefferson
correlation5_jefferson <- cor.test(jefferson_data$low_income, jefferson_data$cash_assist)
correlation5_jefferson
correlation6_jefferson <- cor.test(jefferson_data$cash_assist, jefferson_data$pub_assist)
correlation6_jefferson
```


### Two-Sample T-Test Calculations 


#### <span style="color: blue;">Bay Area T-Test Calculations:</span>  
```{r bay continuous plus grads}
difference1_bay = t.test(total_popE~ majority_grads, 
                    data = bay_data)
difference1_bay
difference2_bay = t.test(low_income ~ majority_grads, 
                    data = bay_data)
difference2_bay
difference3_bay = t.test(pub_assist ~ majority_grads , 
                    data = bay_data)
difference3_bay
difference4_bay = t.test(cash_assist ~ majority_grads , 
                    data = bay_data)
difference4_bay
difference5_bay = t.test(total_popE~ majority_college, 
                    data = bay_data)
difference5_bay
difference6_bay = t.test(low_income ~ majority_college, 
                    data = bay_data)
difference6_bay
difference7_bay = t.test(pub_assist ~ majority_college, 
                    data = bay_data)
difference7_bay
difference8_bay = t.test(cash_assist ~ majority_college, 
                    data = bay_data)
difference8_bay
difference9_bay = t.test(total_popE~ majority_bachelors, 
                    data = bay_data)
difference9_bay
difference10_bay = t.test(low_income ~ majority_bachelors, 
                    data = bay_data)
difference10_bay
difference11_bay = t.test(pub_assist ~ majority_bachelors, 
                    data = bay_data)
difference11_bay
difference12_bay = t.test(cash_assist ~ majority_bachelors, 
                    data = bay_data)
difference12_bay


```

#### <span style="color: green;">Jefferson T-Test Calculations:</span>  

```{r jefferson continuous plus grads}
difference1_jefferson = t.test(total_popE~ majority_grads, 
                    data = jefferson_data)
difference1_jefferson
difference2_jefferson = t.test(low_income ~ majority_grads, 
                    data = jefferson_data)
difference2_jefferson
difference3_jefferson = t.test(pub_assist ~ majority_grads , 
                    data = jefferson_data)
difference3_jefferson
difference4_jefferson = t.test(cash_assist ~ majority_grads , 
                    data = jefferson_data)
difference4_jefferson
difference5_jefferson = t.test(total_popE~ majority_college, 
                    data = jefferson_data)
difference5_jefferson
difference6_jefferson = t.test(low_income ~ majority_college, 
                    data = jefferson_data)
difference6_jefferson
difference7_jefferson = t.test(pub_assist ~ majority_college, 
                    data = jefferson_data)
difference7_jefferson
difference8_jefferson = t.test(cash_assist ~ majority_college, 
                    data = jefferson_data)
difference8_jefferson
difference9_jefferson = t.test(total_popE~ majority_bachelors, 
                    data = jefferson_data)
difference9_jefferson
difference10_jefferson = t.test(low_income ~ majority_bachelors, 
                    data = jefferson_data)
difference10_jefferson
difference11_jefferson = t.test(pub_assist ~ majority_bachelors, 
                    data = jefferson_data)
difference11_jefferson
difference12_jefferson = t.test(cash_assist ~ majority_bachelors, 
                    data = jefferson_data)
difference12_jefferson


```

#### <span style="color: orange;">Bay and Jefferson Combined T-Test Calculations:</span>  

```{r continuous plus region}
difference1_combined = t.test(total_popE ~ region, 
                    data = combined_data)
difference1_combined
difference2_combined = t.test(low_income ~ region, 
                    data = combined_data)
difference2_combined
difference3_combined = t.test(cash_assist ~ region, 
                    data = combined_data)
difference3_combined
difference4_combined = t.test(pub_assist ~ region, 
                    data = combined_data)
difference4_combined



```



### Chi-square test

#### <span style="color: orange;">Bay and Jefferson Combined Chi-square Calculations:</span>  

Because region and county are the categorical variables I'm comparing with my high school graduation variable, I've chosen to perform the tests with the combined data. 

```{r chi-squared 1, class.source = 'fold-show'}
chi1_combined <- chisq.test(combined_data$region, combined_data$majority_grads)
chi1_combined
chi1_combined$observed
chi1_combined$expected
chi2_combined <- chisq.test(combined_data$region, combined_data$majority_college)
chi2_combined
chi2_combined$observed
chi2_combined$expected
chi3_combined <- chisq.test(combined_data$region, combined_data$majority_bachelors)
chi3_combined
chi3_combined$observed
chi3_combined$expected
```


```{r chi-squared 2, class.source = 'fold-show'}
chi4_combined <- chisq.test(combined_data$county, combined_data$majority_grads)
chi4_combined
chi4_combined$observed
chi4_combined$expected
chi5_combined <- chisq.test(combined_data$county, combined_data$majority_college)
chi5_combined
chi5_combined$observed
chi5_combined$expected
chi6_combined <- chisq.test(combined_data$county, combined_data$majority_bachelors)
chi6_combined
chi6_combined$observed
chi6_combined$expected
```

### Analysis of Variance (ANOVA)

####  <span style="color: blue;">Bay Area ANOVA Tests</span>  

##### County and Low Income

```{r anova 1 bay, class.source = 'fold-hide'}
anova1_bay <- aov (low_income ~ county, data = bay_data)
summary(anova1_bay)
```
```{r differences 1 bay, class.source = 'fold-hide'}

differences1_bay <- TukeyHSD(anova1_bay)
summary(differences1_bay)
as_tibble(cbind(pair = row.names(differences1_bay$county),
                differences1_bay$county))
```
##### County and Public Assistance

```{r anova 2 bay, class.source = 'fold-hide'}
anova2_bay <- aov (pub_assist ~ county, data = bay_data)
summary(anova2_bay)
```

```{r differences 2 bay, class.source = 'fold-hide'}

differences2_bay <- TukeyHSD(anova2_bay)
summary(differences2_bay)
as_tibble(cbind(pair = row.names(differences2_bay$county),
                differences2_bay$county))
```
##### County and Cash Assistance

```{r anova 3 bay, class.source = 'fold-hide'}
anova3_bay <- aov (cash_assist ~ county, data = bay_data)
summary(anova3_bay)
```
```{r differences 3 bay, class.source = 'fold-hide'}

differences3_bay <- TukeyHSD(anova3_bay)
summary(differences3_bay)
as_tibble(cbind(pair = row.names(differences3_bay$county),
                differences3_bay$county))
```


####  <span style="color: green;">Jefferson ANOVA Tests</span>  

##### County and Low Income

```{r anova 1 jefferson, class.source = 'fold-hide'}
anova1_jefferson <- aov (low_income ~ county, data = jefferson_data)
summary(anova1_jefferson)
```
```{r differences 1 jefferson, class.source = 'fold-hide'}

differences1_jefferson <- TukeyHSD(anova1_jefferson)
summary(differences1_jefferson)
as_tibble(cbind(pair = row.names(differences1_jefferson$county),
                differences1_jefferson$county))
```
##### County and Public Assistance

```{r anova 2 jefferson, class.source = 'fold-hide'}
anova2_jefferson <- aov (pub_assist ~ county, data = jefferson_data)
summary(anova2_jefferson)
```

```{r differences 2 jefferson, class.source = 'fold-hide'}

differences2_jefferson <- TukeyHSD(anova2_jefferson)
summary(differences2_jefferson)
as_tibble(cbind(pair = row.names(differences2_jefferson$county),
                differences2_jefferson$county))
```
##### County and Cash Assistance

```{r anova 3 jefferson, class.source = 'fold-hide'}
anova3_jefferson <- aov (cash_assist ~ county, data = jefferson_data)
summary(anova3_jefferson)
```
```{r differences 3 jefferson, class.source = 'fold-hide'}

differences3_jefferson <- TukeyHSD(anova3_jefferson)
summary(differences3_jefferson)
as_tibble(cbind(pair = row.names(differences3_jefferson$county),
                differences3_jefferson$county))
```



### Relationship Strengths and Statistical Significance   

**This section contains a summary of the strengths and/or significance of each relationship tested per region. I chose to compile my descriptions here so they can be viewed with easy comparison by region. They're organized by test type and the order they appear within teach test tab, so the reader can reference back to the other tabs if they would like to view the code or calculation output in further detail.** 



#### <span style="color: blue;"> Bay Area Correlation Descriptions</span>  


**Significant relationships are listed below for each region, and defined by the following thresholds:**

.00-.19 = very weak
.20-.39 = weak
.40-.59 = moderate
.60-.79 = strong
.80-1.0 = very strong


1. There is a significant, weak, negative correlation (-.2571) between total population and public assistance in the Bay Area tracts in the dataset. 
2. There is a significant, very strong, positive correlation (.7964) between low income and cash assistance in the Bay Area tracts in the dataset.



#### <span style="color: green;"> Jefferson Correlation Descriptions</span>  

1. There is a significant, moderate, negative correlation (-.4898) between total population and public assistance in the Jefferson Region tracts in the dataset.
2.There is a significant, weak, positive correlation (.2737) between low income and public assistance in the Jefferson Region tracts in the dataset.
3.There is a significant, strong, positive correlation (.6569) between low income and cash assistance in the Jefferson Region tracts in the dataset.
4.There is a significant, weak, positive correlation (.2457) between total population and public assistance in the Jefferson Region tracts in the dataset.



#### <span style="color: orange;"> Bay and Jefferson T-Test Descriptions</span>  

In the Bay Area t-tests, there are significant relationships between the following: 
1. low income and high school graduation rate
2. public assistance and high school graduation rate
3. cash assistance and high school graduation rate
4. low income and bachelor's degree rates
5. public assistance and bachelor's degree rates
6. cash assistance and bachelor's degree rates

In the Jefferson t-tests, there are significant relationships between the following: 
1. public assistance and some college or associates degree rates
2. cash assistance and bachelor's degree rates

#### <span style="color: orange;"> Bay and Jefferson Chi-Square Test Descriptions</span>  

My tests showed that there are statistically significant relationships between county and education rates and region and education rates with the caveat that there are multiple counties where the observed and expected values are less than 5, which compromises the validity of results. 

#### <span style="color: orange;"> Bay and Jefferson ANOVA Test Descriptions</span>  

My ANOVA tests between county and continuous variables found significant relationships between county and all other variables in both regions (though p-values for the Bay Area were universally smaller than in the Jefferson Region. )


### Visualizations, and Comparisons

 
#### Visualizations


##### <span style="color: orange;"> Bay and Jefferson Low Income </span>  

There is a weaker correlation between population and poverty levels in the Bay Area, where there are more people per county, and a greater diversity of areas and tracts, and their respective incomes.  


```{r female pop and poverty bay}
ggplot(bay_data, aes(x = total_popE, y = low_income)) + 
    geom_point(size = 0.5, alpha = 0.25) +
    geom_smooth() +
    theme_bw() +
    scale_x_continuous(name = "Bay  Population") +
    scale_y_continuous(name = "Percent Low Income",
                      breaks = seq(0, 1, by = .1), 
                      labels = paste("",
                                  prettyNum(seq(0, 100, by = 10),
                                            big.mark = ","))) +
    annotate(geom = "text", x = 0, y = .5, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation2_bay$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(correlation2_bay$conf.int[2], digits = 2)),
                         hjust = 0)
```
There is a stronger correlation between population and poverty levels in the Jefferson Region, but considerably fewer datapoints. I included the male population and poverty below because it's interesting to see how the prison population so dramatically skewes the data. 

```{r female pop and poverty jefferson}
ggplot(jefferson_data, aes(x = total_popE, y = low_income)) + 
    geom_point(size = 0.5, alpha = 0.25) +
    geom_smooth() +
    theme_bw() +
    scale_x_continuous(name = "Jefferson Population") +
    scale_y_continuous(name = "Percent Low Income",
                      breaks = seq(0, 1, by = .1), 
                      labels = paste("",
                                  prettyNum(seq(0, 1, by = .1),
                                            big.mark = ","))) +
    annotate(geom = "text", x = 0, y = .5, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(correlation2_jefferson$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(correlation2_jefferson$conf.int[2], digits = 2)),
                         hjust = 0)
```




##### <span style="color: orange;"> Region and Poverty </span>  

One of the most surprising findings this analysis is that Jefferson region has a higher mean number of people per tract living below the poverty line. While more people overall in the Bay Region live below poverty, proportionally more people in the Jefferson region live in poverty.

```{r visualizing region and poverty100}
ggplot(combined_data, aes(x = region, y = low_income)) +
  geom_violin() +
  theme_bw() +
  scale_x_discrete(name = "Region",
                   labels = c("Bay", "Jefferson")) +
   scale_y_continuous(name = "Percentage Low Income by Tract", 
                     breaks = seq(0, 1, by = .1),
                     labels = paste(prettyNum(seq(0, 100, by = 10),
                                              big.mark = ","))) +
  annotate(geom = "text", x = 1.5, y = .5, size = 3,
           label = paste("Difference in means (95% confidence) ",
                         prettyNum(abs(difference2_combined$conf.int[2]), 
                                   digits = 3), " to ",
                         prettyNum(abs(difference2_combined$conf.int[1]), 
                                   digits = 3),
                         sep = "")) +
  annotate(geom = "text", x = 1, y = difference2_combined$estimate[1], size = 3,
           label = paste("Mean (95% confidence)",
                         prettyNum(difference2_combined$estimate[1], digits = 3))) +
  annotate(geom = "text", x = 2, y = difference2_combined$estimate[2], size = 3,
           label = paste("Mean (95% confidence)",
                         prettyNum(difference2_combined$estimate[2], digits = 3)))
```


```{r difference 1 bay plot  }
ggplot(combined_data, aes(x = county, y = low_income)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(name = "County") +
  scale_y_continuous(name = "Population Below 100% Poverty Rate",
                     breaks = seq(0, 1, by = .1),
                     labels = paste(prettyNum(seq(0, 100, by = 10), 
                                              big.mark = ",")))
```


##### <span style="color: green;">Jefferson High School Graduation and Poverty </span>  


```{r visualizing hsgrad and poverty100 jefferson}
ggplot(jefferson_data, aes(x = majority_grads, y =low_income)) +
  geom_violin() +
  theme_bw() +
  scale_x_discrete(name = "Jefferson Tract High School Graduation Rate: Meets or Exceeds 83%",
                   labels = c("Yes", "No")) +
   scale_y_continuous(name = "Jefferson Population Below 100% Poverty by Tract", 
                     breaks = seq(0, 1, by = .1),
                     labels = paste(prettyNum(seq(0, 100, by = 10),
                                              big.mark = ","))) +
  annotate(geom = "text", x = 1.3,  y = .5, size = 4,
           label = paste("Difference in means (95% confidence) ",
                         prettyNum(abs(difference2_combined$conf.int[2]), 
                                   digits = 3), " to ",
                         prettyNum(abs(difference2_combined$conf.int[1]), 
                                   digits = 3),
                         sep = "")) +
  annotate(geom = "text", x = 1, y = difference2_jefferson$estimate[1], size = 3,
           label = paste("Mean (95% confidence)", 
                         prettyNum(difference3_jefferson$estimate[1], digits = 3))) +
  annotate(geom = "text", x = 2, y = difference2_jefferson$estimate[2], size = 3,
           label = paste("Mean (95% confidence)",
                         prettyNum(difference2_jefferson$estimate[2], digits = 3)))
```

