---
title: 'Assignment 2: Distributions and Proportions'
author: "Julia Meinhardt"
date: "9/15/2020"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries}
library(ggplot2)
library(tidyverse)

```

### Reading csv, filtering by region, and attaching. 

```{r load and attach data for Jefferson, message=FALSE}
Jefferson_tract_data <- read_csv("jeffersonbay2018.csv") %>%
  filter(region == "Jefferson")
attach(Jefferson_tract_data)
```

### Assignment Deliverables

For each continuous variable in your dataset calculate:
1.sample mean
2.sample standard deviation
3.95-percent confidence interval for the population mean
4.interquartile range. 

Create a histogram to illustrate the distribution of each variable and describe the distribution in a sentence or two.

For each categorical variable in your dataset calculate:
1.95-percent confidence interval for the proportion of the population in each category


**List of Variables**

1.**female_popE** (continuous)
2.**male_popE**  (continuous)

3.**poverty_100** : total below 100% poverty level (continuous)
4.**poverty_149** : total from 100 to 149% of poverty Level (continuous)
5.**poverty_150** : total at or above 150% of poverty level (continuous)

6.**public_assist** : households receiving public assistance (including ss)  (continuous)
8.**cash_assist** :  households receiving public assistance, cash and food stamps (continuous)
9.**no_assist** (continuous)

10.**majority_hs_grads** (categorical)

11. **county** (categorical)
12. **region** (categorical)


### Summary Stats for Jefferson Region

```{r summary stats female pop Jefferson}

summary(female_popE)
sd(female_popE)
pop_mean <- t.test(female_popE)
pop_mean
```
```{r histogram female population Jefferson}
hist <- ggplot(Jefferson_tract_data, aes(x = female_popE)) +
  geom_histogram() 
hist
```


```{r summary stats male pop Jefferson}

summary(male_popE)
sd(male_popE)
pop_mean <- t.test(male_popE)
pop_mean
```
```{r  histogram male pop Jefferson}
hist <- ggplot(Jefferson_tract_data, aes(x = male_popE)) +
  geom_histogram() 
hist
```

```{r summary stats below 100% poverty Jefferson}
summary(poverty_100E)
sd(poverty_100E)
pop_mean <- t.test(poverty_100E)
pop_mean
```
```{r histogram population below 100% poverty Jefferson}
hist<- ggplot(Jefferson_tract_data, aes(x = poverty_100E)) +
  geom_histogram() 
hist
```

```{r summary 100-149% poverty Jefferson}
summary(poverty_149E)
sd(poverty_149E)
pop_mean <- t.test(poverty_149E)
pop_mean
```
```{r histogram population between 100-149% poverty Jefferson}
hist <- ggplot(Jefferson_tract_data, aes(x = poverty_149E)) +
  geom_histogram() 
hist
```

```{r summary over 150% poverty Jefferson}
summary(poverty_150E)
sd(poverty_150E)
pop_mean <- t.test(poverty_150E)
pop_mean
```
```{r histogram population over 150% poverty Jefferson}
hist <- ggplot(Jefferson_tract_data, aes(x = poverty_150E)) +
  geom_histogram() 
hist
```


```{r summary population receiving public assistance Jefferson}
summary(pub_assistE)
sd(pub_assistE)
pop_mean <- t.test(pub_assistE)
pop_mean
```
```{r histogram population receiving public assistance Jefferson}
hist <- ggplot(Jefferson_tract_data, aes(x = pub_assistE)) +
  geom_histogram() 
hist
```


```{r summary population receiving cash or food stamps public assistance Jefferson}
summary(cash_assistE)
sd(cash_assistE)
pop_mean <- t.test(cash_assistE)
pop_mean
```
```{r histogram population receiving cash or food stamps public assistance Jefferson}
hist <- ggplot(Jefferson_tract_data, aes(x = cash_assistE)) +
  geom_histogram() 
hist
```

```{r summary population not receiving public asssistance}
summary(no_assistE)
sd(no_assistE)
pop_mean <- t.test(no_assistE)
pop_mean
```
```{r histogram population not receiving public assistance }
hist <- ggplot(Jefferson_tract_data, aes(x = no_assistE)) +
  geom_histogram() 
hist
```



###Distribution of Categorical Variables: Jefferson 

```{r t-test majority hs grads Jefferson}
meets_hs_grads <- t.test(majority_hs_grads)
below_hs_grads <- t.test(!majority_hs_grads)
```

```{r hs_grads table}
grads_shares <- tibble(Diplomas = c("Meets Statewise Average",
                                    "Below Statewide Average"), 
                     `Sample` =c(paste(prettyNum(meets_hs_grads$estimate*100,
                                         digits = 3), "%"),
                         paste(prettyNum(below_hs_grads$estimate*100,
                                         digits = 3), "%")), 
                      `95-percent confidence interval` =  
                        c(paste(prettyNum(meets_hs_grads$conf.int[1]*100,
                                          digits = 3), "% to ",
                                prettyNum(meets_hs_grads$conf.int[2]*100,
                                          digits = 3), "%"),
                         paste(prettyNum(below_hs_grads$conf.int[1]*100,
                                          digits = 3), "% to ",
                                prettyNum(below_hs_grads$conf.int[2]*100,
                                          digits = 3), "%")))
knitr::kable(grads_shares, caption = "Census Tracts in Jefferson Region")
```













