---
title: 'Jefferson Region: Distributions and Proportions'
author: "Julia Meinhardt"
date: "9/15/2020"
output: 
    html_document:
      toc: true
      toc_float: true
      highlight:  tango
      code_folding: hide
---

```{r setup,  message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r load libraries, message=FALSE, warning=FALSE}
library(ggplot2)
library(tidyverse)
library(ggthemes)
library(dplyr)
```


```{r load and attach data for Jefferson, message=FALSE, warning=FALSE}
Jefferson_tract_data <- read_csv("jeffersonbay2018.csv") %>%
  filter(region == "Jefferson")
attach(Jefferson_tract_data)
```


## Description and Variable List

This document will explore summary statistics of the following variables:

**List of Variables**

1. Total Population: total_popE =  (continuous)
2. Total below 150 % poverty level: low_income (continuous)
3. Percent of population receiving public assistance (including SSI):  (continuous)
4. Percent of population receiving cash public assistance (including food stamps/SNAP):  (continuous)
5. High School Graduates:  (categorical, >83% meets statewide average)
6. Some college or Associates: (categorical, >29% meets statewide average)
7. Bachelors or higher: (categorical, >33% meets statewide average)
8. County (categorical)
9. Region (categorical)


All calculations are for the  <span style="color: green;">Jefferson Region</span> of my dataset. The Bay Region calculations and distributions are done in a separate file. 

## Summary Stats and Visualizations{.tabset .tabset-fade .tabset-pills}

### Population Summary Stats

**Population is measured as the number of people per tract.** 

```{r summary stats pop Jefferson}

summary(total_popE)
sd(total_popE)
pop_mean1 <- t.test(total_popE)
pop_mean1
```
```{r histogram  population Jefferson}
hist1 <- ggplot(Jefferson_tract_data, aes(x = total_popE)) +
  geom_histogram() 


```

```{r hist1}
hist1+
  theme_solarized() +
  scale_x_continuous(name = "Population Count Per Tract") +
  scale_y_continuous(name = "Total") +
  annotate("text", x = 5000, y = 8, size = 3, 
           label = paste("Population mean (95% confidence):",
                         prettyNum(pop_mean1$conf.int[1], digits = 0),
                         "to",
                         prettyNum(pop_mean1$conf.int[2], digits = 0),
                        
                         "\nSample mean:", 
                         prettyNum(mean(total_popE), digits = 0), 
                         "\nSample median:",
                         prettyNum(median(total_popE), digits = 0),
                         "\nSample standard deviation:",
                         prettyNum(sd(total_popE), digits = 0), 
                         "\nSample interquartile range:",
                         prettyNum(quantile(total_popE)["25%"], digits = 0),
                         "to",
                         prettyNum(quantile(total_popE)["75%"], digits = 0)),
           hjust = 0, vjust = 1) 

```

The outlier population of over 9,000 here is a prison population in Lassen County. 

### Low Income Summary Stats

**Low-income is measured as the percent of the population that is living below 150% poverty per tract.** 

```{r summary stats below 100pct poverty Jefferson}
summary(low_income)
sd(low_income)
pop_mean3 <- t.test(low_income)
pop_mean3
```
```{r histogram population below 100pct poverty Jefferson}
hist3<- ggplot(Jefferson_tract_data, aes(x = low_income)) +
  geom_histogram() 
```

```{r hist 3}
hist3+
  theme_solarized() +
  scale_x_continuous(name = "Percent of Population Below 100% Poverty Rate",
                      breaks = seq(0.0, 0.6, by = 0.1),
                     labels = paste(seq(0, 60, by = 10), "%", sep = ""))+
  scale_y_continuous(name = "Total") +
  annotate("text", x = 0, y = 15, size = 3, 
           label = paste("Population mean (95% confidence):",
                         prettyNum(pop_mean3$conf.int[1], digits = 2),
                         "to",
                         prettyNum(pop_mean3$conf.int[2], digits = 2),
                        
                         "\nSample mean:", 
                         prettyNum(mean(low_income), digits = 2), 
                         "\nSample median:",
                         prettyNum(median(low_income), digits = 2),
                         "\nSample standard deviation:",
                         prettyNum(sd(low_income), digits = 2), 
                         "\nSample interquartile range:",
                         prettyNum(quantile(low_income)["25%"], digits = 2),
                         "to",
                         prettyNum(quantile(low_income)["75%"], digits = 2)),
           hjust = 0, vjust = 1) 

```

### Public Assistance Summary Stats and Distribution

**Public assistance is measured as the percent of the population receiving public assistance (including SSI) per tract.** 

```{r summary population receiving public assistance Jefferson}
summary(pub_assist)
sd(pub_assist)
pop_mean6 <- t.test(pub_assist)
pop_mean6
```
```{r histogram population receiving public assistance Jefferson}
hist6 <- ggplot(Jefferson_tract_data, aes(x = pub_assist)) +
  geom_histogram() 
```

```{r hist6}
hist6+
  theme_solarized() +
  scale_x_continuous(name = "Population Using Public Assistance (Including SSI)", 
                     breaks = seq(0.0, 0.6, by = 0.1),
                     labels = paste(seq(0, 60, by = 10), "%", sep = "")) +
  scale_y_continuous(name = "Total") +
  annotate("text", x = 0, y = 13, size = 3, 
           label = paste("Population mean (95% confidence):",
                         prettyNum(pop_mean6$conf.int[1], digits = 2),
                         "to",
                         prettyNum(pop_mean6$conf.int[2], digits = 2),
                        
                         "\nSample mean:", 
                         prettyNum(mean(pub_assist), digits = ), 
                         "\nSample median:",
                         prettyNum(median(pub_assist), digits = 2),
                         "\nSample standard deviation:",
                         prettyNum(sd(pub_assist), digits = 2), 
                         "\nSample interquartile range:",
                         prettyNum(quantile(pub_assist)["25%"], digits = 2),
                         "to",
                         prettyNum(quantile(pub_assist)["75%"], digits = 2)),
           hjust = 0, vjust = 1) 
```


### Cash Assistance Summary Stats and Distribution

**Cash assistance is measured as the percent of the population receiving cash assistance (including food stamps) per tract.** 

```{r summary population receiving cash or food stamps public assistance Jefferson}
summary(cash_assist)
sd(cash_assist)
pop_mean7 <- t.test(cash_assist)
pop_mean7
```
```{r histogram population receiving cash or food stamps public assistance Jefferson}
hist7 <- ggplot(Jefferson_tract_data, aes(x = cash_assist)) +
  geom_histogram() 
```

```{r hist 7}
hist7+
  theme_solarized() +
  scale_x_continuous(name = "Population Using Public Assistance (Cash and Food Stamps)", 
                     breaks = seq(0.0, 0.5, by = .05),
                     labels = paste(seq(0, 50, by = 5), "%", sep = "")) +
  scale_y_continuous(name = "Total") +
  annotate("text", x = .05, y = 13, size = 3, 
           label = paste("Population mean (95% confidence):",
                         prettyNum(pop_mean7$conf.int[1], digits = 2),
                         "to",
                         prettyNum(pop_mean7$conf.int[2], digits = 2),
                        
                         "\nSample mean:", 
                         prettyNum(mean(cash_assist), digits = 2), 
                         "\nSample median:",
                         prettyNum(median(cash_assist), digits = 2),
                         "\nSample standard deviation:",
                         prettyNum(sd(cash_assist), digits = 2), 
                         "\nSample interquartile range:",
                         prettyNum(quantile(cash_assist)["25%"], digits = 2),
                         "to",
                         prettyNum(quantile(cash_assist)["75%"], digits = 2)),
           hjust = 0, vjust = 1) 
```
Again, the outlier here seems to conicde with the male population outlier causing the histogram to skew right. The spike in this dataset is also seen in several of the histograms above. I'm not sure what this is showing. It could be caused by a tract count with a much higher population count than others in the rural region. 


###  T-Test: High- School Graduates 

**The California statewide high school graduation rate is 83%. High school graduation proportion is categorized as either meeting or exceeding this 83% threshold or being below.**


```{r t-test majority hs grads Jefferson}
meets_hs_grads <- t.test(majority_grads)
below_hs_grads <- t.test(!majority_grads)
```

```{r hs_grads table}
grads_shares <- tibble('High School Graduation Rate' = c("Meets Statewide Average",
                                    "Below Statewide Average"), 
                     `Sample` =c(paste(prettyNum(meets_hs_grads$estimate*100,
                                         digits = 3), "%"),
                         paste(prettyNum(below_hs_grads$estimate*100,
                                         digits = 3), "%")), 
                      `95-percent confidence interval` =  
                        c(paste(prettyNum(meets_hs_grads$conf.int[1]*100,
                                          digits = 3), "% to ",
                                prettyNum(meets_hs_grads$conf.int[2]*100,
                                          digits = 3), "%"),
                         paste(prettyNum(below_hs_grads$conf.int[1]*100,
                                          digits = 3), "% to ",
                                prettyNum(below_hs_grads$conf.int[2]*100,
                                          digits = 3), "%")))
knitr::kable(grads_shares, caption = "Census Tracts in Jefferson Region")
```

### T-Test: Some College or Associates

**The California statewide rate of population with some college or an associates degree is 29%. 'Some college or associates' proportion is categorized as either meeting or exceeding this 29% threshold or being below.** 

```{r t-test majority college  Jefferson}
meets_college <- t.test(majority_college)
below_college <- t.test(!majority_college)
```

```{r college table}
college_shares <- tibble('Some College' = c("Meets Statewide Average",
                                    "Below Statewide Average"), 
                     `Sample` =c(paste(prettyNum(meets_college$estimate*100,
                                         digits = 3), "%"),
                         paste(prettyNum(below_college$estimate*100,
                                         digits = 3), "%")), 
                      `95-percent confidence interval` =  
                        c(paste(prettyNum(meets_college$conf.int[1]*100,
                                          digits = 3), "% to ",
                                prettyNum(meets_college$conf.int[2]*100,
                                          digits = 3), "%"),
                         paste(prettyNum(below_college$conf.int[1]*100,
                                          digits = 3), "% to ",
                                prettyNum(below_college$conf.int[2]*100,
                                          digits = 3), "%")))
knitr::kable(college_shares, caption = "Census Tracts in Jefferson Region")
```

### T-Test: Bachelor's Degree

**The California statewide rate of population with a bachelor's degree is 21%. Bachelor's degree proportions are categorized as either meeting or exceeding this 21% threshold or being below.** 

```{r t-test majority bachelors  Jefferson}
meets_bachelors <- t.test(majority_bachelors)
below_bachelors <- t.test(!majority_bachelors)
```

```{r bachelors table}
bachelors_shares <- tibble('Bachelors Degree'  = c("Meets Statewide Average",
                                    "Below Statewide Average"), 
                     `Sample` =c(paste(prettyNum(meets_bachelors$estimate*100,
                                         digits = 3), "%"),
                         paste(prettyNum(below_bachelors$estimate*100,
                                         digits = 3), "%")), 
                      `95-percent confidence interval` =  
                        c(paste(prettyNum(meets_bachelors$conf.int[1]*100,
                                          digits = 3), "% to ",
                                prettyNum(meets_bachelors$conf.int[2]*100,
                                          digits = 3), "%"),
                         paste(prettyNum(below_bachelors$conf.int[1]*100,
                                          digits = 3), "% to ",
                                prettyNum(below_bachelors$conf.int[2]*100,
                                          digits = 3), "%")))
knitr::kable(bachelors_shares, caption = "Census Tracts in Jefferson Region")


```









```{r save as .csv,  message=FALSE, warning=FALSE}

write_csv(Jefferson_tract_data, "jeffersontractdata2018.csv")

```















