---
title: "Regression Part 2"
author: "Julia Meinhardt"
date: "10/3/2020"
output: html_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(jtools)
library(interactions)
library(knitr)
```

Estimate at least one more version of the regression you estimated in Assignment 4 by transforming one or more of the independent variables and including at least one interaction term. Which version of the model best explains the observed variation in your dependent variable?


### Regression Model: Transforming Variables 
```{r}

combined_mutated <- Jefferson_Bay <- Jefferson_Bay %>%
  mutate(low_income = case_when(pub_assist < .36 ~ "1st Quartile",
                                       pub_assist > .36  ~ "Average"))

combined_mutated <- lm(cash_assist ~ total_popE + pub_assist +low_income + majority_grads + majority_college +  majority_bachelors + region,
            data = Jefferson_Bay)

summary(combined_mutated)
```
What does it mean if my r value goes down but my significance goes up??

TO DO: Add interaction Terms for Region and Visualizw 








### Scraps


```{r}

summary(Jefferson_Bay$cash_assist)
summary(Jefferson_Bay$pub_assist)

```





```{r}
Bay_tract_data <- Bay_tract_data %>%
  mutate(mostly_low_income = case_when(pub_assist < .36 ~ "1st Quartile",
                                       pub_assist > .36  ~ "Average"))

model_transformed <- lm(low_income ~ total_popE + pub_assist + cash_assist + majority_grads + majority_college +  majority_bachelors,
            data = Bay_tract_data)

summary(model_transformed)
```



