---
title: "Assignment 1: Assembling a Dataframe"
author: "Julia Meinhardt"
date: "9/10/2020"
output: html_document
---

## Area of Study

The dataset I compiled for this exercise will be used to explore two different different regions of Northern California: the Bay Area and an informal region sometimes referred to as "Jefferson." I'm interested in residents' poverty levels and use of public assistance both within each respective region and between the two regions for comparison. 

*I utilized Kristy's code and methods extensively in this dataset! I both accessed her code on GitHub and asked her questions personally when I was stumped. Her code and assistance were essential in organizing my data by county (and I directly used her Bay Area counties code to combine the tracts into counties) and region and turning them into a categorical variables, as well as binding my datasets together into one dataframe.*

## Loading and Choosing Variables

First, I loaded the necessary libraries. 

```{r load libraries, message = FALSE,results = 'hide'  }

library(tidycensus)
library(tidyverse)
library(dplyr)
library(ggplot2)
```

Next, I loaded area-level data from the 5-Year ACS.

```{r load area-level data}
area_vars_2018 <- load_variables (2018, "acs5")

```


I chose the following variables to include in my dataset: 

*The number of variabes is larger than I initially expected, but after exploring the variables in each dataset, I determined that a closer look into levels of poverty and types of government assistance would enable me to explore my area of interest with greater nuance and detail.*


1.Total Population: total_pop =  (continuous)
2.Total below 150 % poverty level: low_income (continuous)
3.Percent of population receiving public assistance (including SSI):  (continuous)
4.Percent of population receiving cash public assistance (including food stamps/SNAP):  (continuous)
5.High School Graduates:  (mutated into categorical, >83% meets statewide average)
6.Some college or Associates: (>29% meets statewide average)
7.Bachelors or higher: (>33 meets statewide average)
8. County (categorical)
9. Region (categorical)


## Loading, Joining, and Creating Variables

CHANGES: The only thing that has changed in lines 69-273 is in lines 75-82. I eliminated male/female, poverty 150, no assist, and added some college and bachelors. The problem I'm running into centers around the hs_grads variable. 


 OLD CODE: population_vars <- c( total_pop = "B00001_001")

              male_pop =    "B01001_002",
              female_pop =  "B01001_026", 
              
              poverty_100 = "B06012_002",
              poverty_149 = "B06012_003",
              poverty_150 = "B06012_004",
              
              pub_assist =  "B19057_001",
              cash_assist = "B19058_002",
              no_assist =   "B19057_003",
              
              hs_grads =    "B06009_003")    
              
              
``` {r load variables for dataframe, message=FALSE, include = TRUE, results = 'hide'}

population_vars <- c( total_pop = "B02001_001", 
              poverty_100 = "B06012_002",
              poverty_149 = "B06012_003",
              pub_assist =  "B19057_001",
              cash_assist = "B19058_002",
              hs_grads =    "B06009_003",
              some_college = "B06009_004",
              bachelors = "B06009_005")

tractLassen2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Lassen",
              year = 2018, 
              survey = "acs5")

tractLassen2018$county <- rep("Lassen", times = length(tractLassen2018$total_popE))
tractLassen2018$region <- rep("Jefferson", times = length(tractLassen2018$total_popE))


tractModoc2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Modoc",
              year = 2018, 
              survey = "acs5")

tractModoc2018$county <- rep("Modoc", times = length(tractModoc2018$total_popE))
tractModoc2018$region <- rep("Jefferson", times = length(tractModoc2018$total_popE))


tractPlumas2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Plumas",
              year = 2018, 
              survey = "acs5")

tractPlumas2018$county <- rep("Plumas", times = length(tractPlumas2018$total_popE))
tractPlumas2018$region <- rep("Jefferson", times = length(tractPlumas2018$total_popE))


tractShasta2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Shasta",
              year = 2018, 
              survey = "acs5")

tractShasta2018$county <- rep("Shasta", times = length(tractShasta2018$total_popE))
tractShasta2018$region <- rep("Jefferson", times = length(tractShasta2018$total_popE))


tractTrinity2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Trinity",
              year = 2018, 
              survey = "acs5")

tractTrinity2018$county <- rep("Trinity", times = length(tractTrinity2018$total_popE))
tractTrinity2018$region <- rep("Jefferson", times = length(tractTrinity2018$total_popE))


tractSiskiyou2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Siskiyou",
              year = 2018, 
              survey = "acs5")

tractSiskiyou2018$county <- rep("Siskiyou", times = length(tractSiskiyou2018$total_popE))
tractSiskiyou2018$region <- rep("Jefferson", times = length(tractSiskiyou2018$total_popE))


tractDelNorte2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Del Norte",
              year = 2018, 
              survey = "acs5")

tractDelNorte2018$county <- rep("Del Norte", times = length(tractDelNorte2018$total_popE))
tractDelNorte2018$region <- rep("Jefferson", times = length(tractDelNorte2018$total_popE))


tractAlameda2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Alameda",
              year = 2018, 
              survey = "acs5")

tractAlameda2018$county <- rep("Alameda", times = length(tractAlameda2018$total_popE))
tractAlameda2018$region <- rep("Bay", times = length(tractAlameda2018$total_popE))


tractContraCosta2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Contra Costa",
              year = 2018, 
              survey = "acs5")

tractContraCosta2018$county <- rep("ContraCosta", times = length(tractContraCosta2018$total_popE))
tractContraCosta2018$region <- rep("Bay", times = length(tractContraCosta2018$total_popE))


tractMarin2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Marin",
              year = 2018, 
              survey = "acs5")

tractMarin2018$county <- rep("Marin", times = length(tractMarin2018$total_popE))
tractMarin2018$region <- rep("Bay", times = length(tractMarin2018$total_popE))


tractNapa2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Napa",
              year = 2018, 
              survey = "acs5")

tractNapa2018$county <- rep("Napa", times = length(tractNapa2018$total_popE))
tractNapa2018$region <- rep("Bay", times = length(tractNapa2018$total_popE))


tractSanFrancisco2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "San Francisco",
              year = 2018, 
              survey = "acs5")

tractSanFrancisco2018$county <- rep("San Francisco", times = length(tractSanFrancisco2018$total_popE))
tractSanFrancisco2018$region <- rep("Bay", times = length(tractSanFrancisco2018$total_popE))


tractSanMateo2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "San Mateo",
              year = 2018, 
              survey = "acs5")

tractSanMateo2018$county <- rep("San Mateo", times = length(tractSanMateo2018$total_popE))
tractSanMateo2018$region <- rep("Bay", times = length(tractSanMateo2018$total_popE))


tractSantaClara2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Santa Clara",
              year = 2018, 
              survey = "acs5")

tractSantaClara2018$county <- rep("Santa Clara", times = length(tractSantaClara2018$total_popE))
tractSantaClara2018$region <- rep("Bay", times = length(tractSantaClara2018$total_popE))


tractSolano2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Solano",
              year = 2018, 
              survey = "acs5")

tractSolano2018$county <- rep("Solano", times = length(tractSolano2018$total_popE))
tractSolano2018$region <- rep("Bay", times = length(tractSolano2018$total_popE))


tractSonoma2018 <- get_acs(geography= "tract", 
              variables = population_vars, 
              output = "wide", 
              state = "CA",
              county = "Sonoma",
              year = 2018, 
              survey = "acs5")

tractSonoma2018$county <- rep("Sonoma", times = length(tractSonoma2018$total_popE))
tractSonoma2018$region <- rep("Bay", times = length(tractSonoma2018$total_popE))

Jefferson_Bay <- rbind(tractModoc2018, tractLassen2018, tractPlumas2018, tractShasta2018, tractTrinity2018, tractSiskiyou2018, tractDelNorte2018, tractAlameda2018, tractContraCosta2018, tractMarin2018, tractNapa2018, tractSanFrancisco2018, tractSanMateo2018, tractSantaClara2018, tractSolano2018, tractSonoma2018)

```

PROBLEMS HERE: 

After running the above chunk, you can see that the number of Hs grads/per tract (majority_grads)is coming up very low, so when I run my mutations below, it's reading all of my observations as FALSE because they're below the 0.83 threshold. This is in contrast to my first time through this dataset, where roughly 90ish % of tracts were reading as having >83% population having high school diplomas. 

The majority_college and majority_bachelors variables seem to be calculating just fine. 

```{r mutate and select}
Jefferson_Bay <- Jefferson_Bay %>%
  mutate(majority_grads = hs_gradsE/total_popE > 0.83,
         majority_college = some_collegeE/total_popE > 0.29,
         majority_bachelors = bachelorsE/total_popE > 0.33,
         cash_assist = cash_assistE/total_popE,
         pub_assist = pub_assistE/total_popE,
         low_income = (poverty_100E + poverty_149E)/total_popE) %>%
  filter(total_popE != 0)%>%
  select (NAME, low_income, total_popE, pub_assist, cash_assist, majority_grads, majority_college, majority_bachelors, region, county)
```

A sample of the assembled dataframe is shown below. 

```{r}
Jefferson_Bay
```
Lastly, I saved the dataframeas a .csv for future use. 

```{r save as .csv}

write_csv(Jefferson_Bay, "jeffersonbay2018.csv")

```


write_csv(Bay2018, "baytractdata2018.csv")
```

## Summary

1. Unit of analysis: county (by census tract)
2. Definition of sample population: all residents of 16 counties in California 
3. Variables: population (male, female, total), poverty level (100%, 100-149%, 150%) public assistance (use, type) region, county, education 
4. The number of observations (rows) in dataset: 1,683


